First of all, I am so grateful to my main supervisor Dr. Matthew Parry for offering me this opportunity to take part in the collaborative project with the Physics department, and to work with him and other talented researchers. His extensive knowledge of statistics and physics leads me in the right way and encourages me throughout this difficult project. His passionate at exploring the unknown world of science motivated me and will motivate me to go further and deeper in the area of developing and implementing statistical models.  

Secondly, I would like to appreciate my co-supervisor Professor David Bryant for his great help in my research progress and careful reading of the manuscript. He offered me extraordinary supports during the time when Matthew was overseas. His guidance and expert advice inspire me and lit my way in finding a different method. 

A special gratitude goes out to the Ministry of Business, Innovation and Employment for supporting my research. This project would have been impossible without their generous fundings. Additionally, many thanks to TracMap company for providing GPS data set, which is the base stone of all the work. 

Moreover, I want to thank all the support staff at Mathematics and Statistics Department for their great job, including Lenette, Marguerite, Leanne, Chris and Greg, and give many thanks to my office mates and friends, including but not limited to Paula, Chuen Yan, Johannes, Vivian, Juan, Bob and his family, Freddy and his family, for their friendship and accompanying me is the last three years. We had enjoyable and unforgettable experiences together. 

Finally, last but by no means least, I will thank my parents for their wise counsel and encouragements to me for studying overseas, and my beloved wife Yingying for all her love and support, and my daughter Anyang, who was born in my second year of study. Being her father is the greatest accomplishment of my life.

Thanks for all your encouragements! 

